Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-04-17 manuscript/manuscript.tex
Removed unnecesary equation from manuscript
Modified exponential-delta.pdf figure and added in the manuscript

\begin{equation}
    \text{max}\{ \Delta \rho(r') \} \frac{L_r^\text{sm}}{L_r} > \delta,
    \label{eq:delta-density}
\end{equation}
\begin{equation}
    \text{max}\{ \Delta \rho(r') \} = \Delta \rho(R_\text{max}),
\end{equation}

@@

\begin{figure*}
\centering
\includegraphics[width=0.9\linewidth]{figures/linear-D.pdf}
\caption{
    Differences between the gravity fields generated by the numerical 
    model and the analytical solution for (a) a thin spherical shell of 
    1km of thickness and (b) a thick spherical shell of 35km of thickness,
    both with the linear density defined on equation 
    \ref{eq:density-linear}.
    The computations were performed on the four grids described in 
    Table \ref{tab:grids}, for different values of the distance-size 
    ratio $D$ and without applying the density-based discretization 
    algorithm.
    If the difference is less than 0.1\%, we consider that the model 
    has achieved an acceptable accuracy for the corresponding value of 
    $D$.
    }
\label{fig:D-linear}
\end{figure*}

@@

>>>>

\begin{equation}
    \text{max}\{ \Delta \rho(r') \} \frac{L_r^\text{sm}}{L_r} > \delta,
    \label{eq:delta-density}
\end{equation}

@@

\begin{figure*}
\centering
\includegraphics[width=0.9\linewidth]{figures/linear-D.pdf}
\caption{
    Differences between the gravity fields generated by the numerical 
    model and the analytical solution for (a) a thin spherical shell of 
    1km of thickness and (b) a thick spherical shell of 35km of thickness,
    both with the linear density defined on equation 
    \ref{eq:density-linear}.
    The computations were performed on the four grids described in 
    Table \ref{tab:grids}, for different values of the distance-size 
    ratio $D$.
    Due to the linearity of the density function the density-based 
    discretization algorithm is not applied.
    If the difference is less than 0.1\%, we consider that the model 
    has achieved an acceptable accuracy for the corresponding value of 
    $D$.
    }
\label{fig:D-linear}
\end{figure*}

@@

Taking this conclusion into account we have determined the $\delta$ 
value needed to ensure an acceptable accuracy through a comparison 
between the numerical approximation and the analytical solution for a 
spherical shell with an exponential density.
In this specific case we proposed an exponential density function  that 
assumes the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's 
outer and inner radii, respectively. It can be defined as follows:

\noindent where $R$ is the mean Earth radius and $b$ is a constant 
that determines the variability of the function: a low value of $b$ 
increases the maximum slope of the density function.

In order to perform this comparison we must compute the analytical 
solution of the gravitational fields generated by this spherical shell 
and define a tesseroid model to numerically approximate this fields.

Firstly, the analytical solution can be computed as the sum of the 
fields generated by a spherical shell with constant density $\rho(r') = 
C$ and the one determined in equation \ref{eq:shell-pot-exp}.

Secondly, we proposed a thin and a thick spherical shells with 1km and 
35km of thickness, respectively, modelled by a mesh of $30^\circ \times 
30^\circ$ tesseroids.
We have assigned the exponential density function proposed on equation 
\ref{eq:density-exp} to each spherical shell model, choosing few values 
of the constant $b$ in order to test the numerical approximation for 
different density variations.

Finally, we computed the gravitational potential ($V$), the 
vertical component of its gradient ($g_z$) and the $g_zz$ component of 
the Marussi tensor on the big grid (see Table \ref{tab:grids}) by 
keeping the distance-size ratio equal to the $D$ values obtained in the 
density case for its corresponding gravitational field and exploring 
different values of $\delta$.
We then computed the percentage difference between this numerical 
results and the analytical solutions.

On Figures \ref{fig:delta-exponential}(a) and 
\ref{fig:delta-exponential}(b) we show the density functions used in 
the comparison and the resulting percentage difference for each 
gravitational field.
The former shows the results for the thin spherical shell, while the 
later shows the ones corresponding to the thick shell.
From these results we can conclude that for the potential and $g_z$ 
computation, a $\delta = 0.3$ is needed in order to achieve an 
acceptable numerical error, while a $\delta = 0.02$ is required for the 
$g_{zz}$ Marussi tensor component.

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{figures/exponential-delta.pdf}
\caption{
    Density functions used in the model comparison and the differences 
    between the gravitational fields generated by the numerical model 
    and the analytical solution for (a) a thin spherical shell of 1km 
    of thickness and (b) a thick spherical shell of 35km of thickness, 
    both with the exponential density defined on equation 
    \ref{eq:density-exp}.
    A few values of the constant $b$ were used in order to test the 
    numerical model for different density variations.
    The computations were performed on the big grid described in 
    Table \ref{tab:grids}, with fixed values of the distance-size ratio 
    $D$ obtained for in the linear density case ($D=1, 2, 8$ for the 
    potential, gradient components and tensor components, respectively)
    and different values of the $\delta$ ratio.
    If the difference is less than 0.1\%, we consider that the model 
    has achieved an acceptable accuracy for the corresponding value of 
    $D$.
    }
\label{fig:delta-exponential}
\end{figure*}
